---
title: "Supporting information"
author: "Tom Ellis"
date: "June 14, 2017"
output:
  pdf_document:
    fig_caption: yes
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
library("qtl", quietly = T)
library(qtltools)
```

This Rmarkdown file contains the files necessary to recreate the results presented in the manuscript "Components of fecundity and local adaptation in *Arabidopsis thaliana*" by Tom Ellis and Jon Ågren. If you are reading a PDF, HTML or MS Word version of this document, see the original `.Rmd` file for the full R source code. See also the file `main_results.Rmd` for code to generate the data in the tables.

# Supporting tables

```{r tab:mann-whitney-U}
library(knitr)
utable <- read.csv('../tables/mannwhitneyu_parental_means.csv')
kable(utable, digits=c(1,1,4,1,4,1,4), row.names = F,
             col.names = c("experiment", "U (fruit)", "p-value (fruit)","U (seeds)", "p-value (seeds)","U (fecundity)", "p-value (fecundity)"))
```

Table S1: Mann-Whitney-U tests on differences between parental lines. Columns show test statistics and p-values for fruit number per plant, seed number per fruit and seed number per plant.

```{r rqtl data import, cache=T, cache.lazy=T, message=F}
# Import R/qtl data on seed number, mass and total estimated fitness (seed number * total fruit number).
numbqtl <- read.cross("csv", "", "../qtl_mapping/rqtl_numb.csv", na.strings=NA, genotypes = c("a", "b"), alleles=c("I","S"))
tofuqtl <- read.cross("csv", "", "../qtl_mapping/rqtl_tofu.csv", na.strings=NA, genotypes = c("a", "b"), alleles=c("I","S"))
frutqtl <- read.cross("csv", "", "../qtl_mapping/rqtl_frut.csv", na.strings=NA, genotypes = c("a", "b"), alleles=c("I","S"))
# state that this is RIL data from a selfer
numbqtl <- convert2riself(numbqtl)
tofuqtl <- convert2riself(tofuqtl)
frutqtl <- convert2riself(frutqtl)
# get genotype probabilities for each rqtl object
numbqtl <- calc.genoprob(numbqtl, step=2, error.prob=.0001, map.function="kosambi",stepwidth="max", off.end=0)
tofuqtl <- calc.genoprob(tofuqtl, step=2, error.prob=.0001, map.function="kosambi",stepwidth="max", off.end=0)
frutqtl <- calc.genoprob(frutqtl, step=2, error.prob=.0001, map.function="kosambi",stepwidth="max", off.end=0)
# coerce values based on <3 values to NA
for(c in 2:11) numbqtl$pheno[,c][numbqtl$pheno[,c+10] <3] <- NA

# import Rdata objects containing stepwise QTL fitting results.
load('qtl_mapping/stepwise_frut.Rdata')
load('qtl_mapping/stepwise_numb.Rdata')
load('qtl_mapping/stepwise_tofu.Rdata')
```

```{r drop1 analyses, cache=T, cache.lazy=TRUE}
# empty lists for the model fits.
numb_fit <- mass_fit <- tofu_fit <- frut_fit <- vector('list', 10) 
names(numb_fit) <- names(mass_fit) <- names(tofu_fit) <- names(frut_fit) <- names(numb_stepwise)

# drop one analyses for each site-year combination.
for(y in 1:10){
  if(y != 9){ # for seed number there were no qtl in 2014, so only loop up to eight.
    numb_fit[[y]] <- fitqtl(numbqtl, pheno.col=y+1,  qtl=numb_stepwise[[y]], formula=formula(numb_stepwise[[y]]),
                            method="hk", model="normal", covar=NULL, get.ests=T, dropone=T)
    frut_fit[[y]] <- fitqtl(frutqtl, pheno.col=y+1,  qtl=frut_stepwise[[y]], formula=formula(frut_stepwise[[y]]),
                            method="hk", model="normal", covar=NULL, get.ests=T, dropone=T)
    tofu_fit[[y]] <- fitqtl(tofuqtl, pheno.col=y+1,  qtl=tofu_stepwise[[y]], formula=formula(tofu_stepwise[[y]]),
                            method="hk", model="normal", covar=NULL, get.ests=T, dropone=T)
  }
}

```

```{r create tab:all-qtl-models}
# create a summary table of all QTL identified
# the clustering function from qtltools can pull out the relevant information easily.
clusters <- list(numb = cluster_qtl(1:5, numb_stepwise[-9], numb_fit[-9]),
                 tofu = cluster_qtl(1:5, tofu_stepwise[-9], tofu_fit[-9]),
                 frut = cluster_qtl(1:5, frut_stepwise[-9], frut_fit[-9]))
# pull out the data table listing all loci.
allqtl_frut <- clusters$frut$full.list
allqtl_numb <- clusters$numb$full.list
allqtl_tofu <- clusters$tofu$full.list
# the data tables have experiment listed by number
# to give these a meaning we need to insert values for site an year.
# first a reference table
exptable <- data.frame(site = c(rep("Italy",5), rep("Sweden",4)),
                       year = c(2010:2014, 2010:2012, 2014),
                       index= 1:9)
# Insert site and year information for fruit number
ix <- match(allqtl_frut$experiment, exptable$index)
allqtl_frut$site <- exptable$site[ix]
allqtl_frut$year <- exptable$year[ix]
# Insert site and year information for seed number
ix <- match(allqtl_numb$experiment, exptable$index)
allqtl_numb$site <- exptable$site[ix]
allqtl_numb$year <- exptable$year[ix]
# Insert site and year information for fecundity
ix <- match(allqtl_tofu$experiment, exptable$index)
allqtl_tofu$site <- exptable$site[ix]
allqtl_tofu$year <- exptable$year[ix]

# string term for 95% Bayesian credible intervals
allqtl_frut$bayesian_CIs <- paste(round(allqtl_frut$min_bayesint, 1), round(allqtl_frut$max_bayesint, 1), sep="-")
allqtl_numb$bayesian_CIs <- paste(round(allqtl_numb$min_bayesint, 1), round(allqtl_numb$max_bayesint, 1), sep="-")
allqtl_tofu$bayesian_CIs <- paste(round(allqtl_tofu$min_bayesint, 1), round(allqtl_tofu$max_bayesint, 1), sep="-")
# sort by site and year
allqtl_frut <- allqtl_frut[order(allqtl_frut$site, allqtl_frut$year),]
allqtl_numb <- allqtl_numb[order(allqtl_numb$site, allqtl_numb$year),]
allqtl_tofu <- allqtl_tofu[order(allqtl_tofu$site, allqtl_tofu$year),]
# pull out the variables of interest
jx <- c("site", "year", "chr", "ML_bayesint", "bayesian_CIs", "LOD", "effect_sizes", "PVE")
allqtl_frut <- allqtl_frut[, jx]
allqtl_numb <- allqtl_numb[, jx]
allqtl_tofu <- allqtl_tofu[, jx]
# round variables
for(c in c(4,6,7,8)){
  allqtl_frut[,c] <- round(allqtl_frut[,c], 1)
  allqtl_numb[,c] <- round(allqtl_numb[,c], 1)
  allqtl_tofu[,c] <- round(allqtl_tofu[,c], 1)
}
# write to disk
write.csv(allqtl_frut, file='tables/all_qtl_frut.csv', row.names = F)
write.csv(allqtl_frut, file='tables/all_qtl_numb.csv', row.names = F)
write.csv(allqtl_frut, file='tables/all_qtl_tofu.csv', row.names = F)
```

```{r tab:individual-frut-qtl}
cnames <- c("site", "year", "chr", "pos (cM)", "Bayesian 95% CIs", "LOD", "effect size", "PVE")
kable(allqtl_frut, row.names = F, col.names = cnames)
```

Table S2: QTL for fruit number per plant identifed in individual experiments. Columns list site, year, chromosome, maximum-likelihood marker position, Bayesian 95% credible intervals for QTL position, LOD score, effect size of the Swedish allele, and proportion of the phenotypic variance explained.

```{r tab:individual-numb-qtl}
cnames <- c("site", "year", "chr", "pos (cM)", "Bayesian 95% CIs", "LOD", "effect size", "PVE")
kable(allqtl_numb, row.names = F, col.names = cnames)
```

Table S3: QTL for seed number per fruit identifed in individual experiments. Columns list site, year, chromosome, maximum-likelihood marker position, Bayesian 95% credible intervals for QTL position, LOD score, effect size of the Swedish allele, and proportion of the phenotypic variance explained.

```{r tab:individual-tofu-qtl}
cnames <- c("site", "year", "chr", "pos (cM)", "Bayesian 95% CIs", "LOD", "effect size", "PVE")
kable(allqtl_tofu, row.names = F, col.names = cnames)

```

Table S4: QTL for seed number per plant identifed in individual experiments. Columns list site, year, chromosome, maximum-likelihood marker position, Bayesian 95% credible intervals for QTL position, LOD score, effect size of the Swedish allele, and proportion of the phenotypic variance explained.

```{r cluster QTL}
# cluster QTL based on credible intervals.
# In fact we did this before, but I will repeat the code here for context.
library(qtltools)
# clustering for QTL at both sites
clusters <- list(frut = cluster_qtl(1:5, frut_stepwise[-9], frut_fit[-9], 15.2),
                 numb = cluster_qtl(1:5, numb_stepwise[-9], numb_fit[-9], 15.2),
                 tofu = cluster_qtl(1:5, tofu_stepwise[-9], tofu_fit[-9], 15.2))
cluster_bysite <- list(frut_it = cluster_qtl(1:5, frut_stepwise[1:5 ],        frut_fit[1:5 ], 15.2),
                       numb_it = cluster_qtl(1:5, numb_stepwise[1:5 ],        numb_fit[1:5 ], 15.2),
                       tofu_it = cluster_qtl(1:5, tofu_stepwise[1:5 ],        tofu_fit[1:5 ], 15.2),
                       frut_sw = cluster_qtl(1:5, frut_stepwise[c(6,7,8,10)], frut_fit[c(6,7,8,10)], 15.2),
                       numb_sw = cluster_qtl(1:5, numb_stepwise[c(6,7,8,10)], numb_fit[c(6,7,8,10)], 15.2),
                       tofu_sw = cluster_qtl(1:5, tofu_stepwise[c(6,7,8,10)], tofu_fit[c(6,7,8,10)], 15.2))
```

```{r cluster summary fruit number}
library(qtltools)
nq_it <- cluster_qtl_summary(clusters$frut$full.list[clusters$frut$full.list$experiment %in% 1:5,])
nq_sw <- cluster_qtl_summary(clusters$frut$full.list[clusters$frut$full.list$experiment %in% 6:10,])

nq_it <- nq_it[match(clusters$frut$summary$QTL_id, nq_it$QTL_id),]
nq_sw <- nq_sw[match(clusters$frut$summary$QTL_id, nq_sw$QTL_id),]

nq_it$n.qtl[is.na(nq_it$n.qtl)] <- 0
nq_sw$n.qtl[is.na(nq_sw$n.qtl)] <- 0

knitr::kable(
  data.frame(QTL=paste("F", row.names(clusters$frut$summary), sep=""),
             clusters$frut$summary$chr,
             clusters$frut$summary$pos_mean,
             nq_it$n.qtl,
             nq_sw$n.qtl),
  digits=2, col.names = c("QTL", "chromosome", "weighted mean position (cM)", "frequency in Italy", "frequency in Sweden"), row.names = F)
```

Table S5: QTL clusters for fruit number per plant.

```{r cluster summary seed number}
nq_it <- cluster_qtl_summary(clusters$numb$full.list[clusters$numb$full.list$experiment %in% 1:5,])
nq_sw <- cluster_qtl_summary(clusters$numb$full.list[clusters$numb$full.list$experiment %in% 6:10,])

nq_it <- nq_it[match(clusters$numb$summary$QTL_id, nq_it$QTL_id),]
nq_sw <- nq_sw[match(clusters$numb$summary$QTL_id, nq_sw$QTL_id),]

nq_it$n.qtl[is.na(nq_it$n.qtl)] <- 0
nq_sw$n.qtl[is.na(nq_sw$n.qtl)] <- 0

knitr::kable(
  data.frame(QTL=paste("N", row.names(clusters$numb$summary), sep=""),
             clusters$numb$summary$chr,
             clusters$numb$summary$pos_mean,
             nq_it$n.qtl,
             nq_sw$n.qtl),
  digits=2, col.names = c("QTL", "chromosome", "weighted mean position (cM)", "frequency in Italy", "frequency in Sweden"), row.names = F)
```

Table S6: QTL clusters for seed number per fruit.

```{r cluster summary total fecundity}
nq_it <- cluster_qtl_summary(clusters$tofu$full.list[clusters$tofu$full.list$experiment %in% 1:5,])
nq_sw <- cluster_qtl_summary(clusters$tofu$full.list[clusters$tofu$full.list$experiment %in% 6:10,])

nq_it <- nq_it[match(clusters$tofu$summary$QTL_id, nq_it$QTL_id),]
nq_sw <- nq_sw[match(clusters$tofu$summary$QTL_id, nq_sw$QTL_id),]

nq_it$n.qtl[is.na(nq_it$n.qtl)] <- 0
nq_sw$n.qtl[is.na(nq_sw$n.qtl)] <- 0

knitr::kable(
  data.frame(QTL=paste("T", row.names(clusters$tofu$summary), sep=""),
             clusters$tofu$summary$chr,
             clusters$tofu$summary$pos_mean,
             nq_it$n.qtl,
             nq_sw$n.qtl),
  digits=2, col.names = c("QTL", "chromosome", "weighted mean position (cM)", "frequency in Italy", "frequency in Sweden"), row.names = F)
```

Table S7: QTL clusters for seed number per plant.

# Supporting figures

```{r fig:ind-years-fruit, fig.width=21/2.56}
mloc <- read.table('../data_files/genetic_map_agren_etal_2013.csv', sep = '\t', header=T)
# pdf('../figures/QTL_positions_frut.pdf', width=23/2.56, height = 18/2.56, family='Times')
par(mfrow=c(1,1), mar=c(0,3,0,0)+0.1, oma=c(1,0,0,0), family='Times')
# set up plot
frut_plot <- plotQTL(1:5, mloc, nlanes = 10, left_gap = 4, right_gap = 0)
plotQTL_base(frut_plot, by=20)
text(33, -92, "Chromosome")
# import fitness boxes and plot.
boxes <- read.csv("../data_files/fitness_boxes.csv")
boxes$track <- match(boxes$chr, colnames(frut_plot$lane_margins)) # vector denoting which track to plot to.
for(b in 1:nrow(boxes)) plotQTL_box(frut_plot, boxes$track[b], boxes$upper[b], boxes$lower[b])
# Lanes and labels
plotQTL_lanes(frut_plot)
segments(frut_plot$lane_margins[6,], 0, frut_plot$lane_margins[6,], -frut_plot$track_lengths)
text(frut_plot$lane_centres,     3, 2010:2014, col="gray50", cex=0.7, srt=90)
text(frut_plot$lane_centres[3,], 8, "Italy",   col='red3',   cex=0.8, adj = c(0.5,0))
text(frut_plot$lane_centres[8,], 8, "Sweden",  col='blue4',  cex=0.8, adj = c(0.5,0))
# plot QTL for each site-year combination
for(y in 1:5){
             plotQTL_qtl(frut_plot, frut_stepwise[[y  ]], frut_fit[[y  ]], lane = y,   col = 'red3',  fill_cutoff = 15.2, lwd=2)
  if(y != 4) plotQTL_qtl(frut_plot, frut_stepwise[[y+5]], frut_fit[[y+5]], lane = y+5, col = 'blue4', fill_cutoff = 15.2, lwd=2)
}

# dev.off()
```

Figure S1: QTL for fruit number per plant in each year in Italy (red) and Sweden (blue). Triangle show the maximum-likelihood position of the QTL and bars indicate the extent of 95% Bayesian credible intervals. The direction of the arrow indicates the direction of effect of the Swedish allele on the phenotype. Open triangles are 'poorly defined' QTL with credible intervals greater than 15.2cM.

```{r fig:ind-years-seeds, fig.width=21/2.56}
library(qtl)
library(qtltools)
mloc <- read.table('../data_files/genetic_map_agren_etal_2013.csv', sep = '\t', header=T)

# pdf('../figures/QTL_positions_number.pdf', width=23/2.56, height = 18/2.56, family='Times')
par(mfrow=c(1,1), mar=c(0,3,0,0)+0.1, oma=c(1,0,0,0), family='Times')
# set up plot
numb_plot <- plotQTL(1:5, mloc, nlanes = 10, left_gap = 4, right_gap = 0)
plotQTL_base(numb_plot, by=20)
text(33, -92, "Chromosome")
# import fitness boxes and plot.
boxes <- read.csv("../data_files/fitness_boxes.csv")
boxes$track <- match(boxes$chr, colnames(numb_plot$lane_margins)) # vector denoting which track to plot to.
for(b in 1:nrow(boxes)) plotQTL_box(numb_plot, boxes$track[b], boxes$upper[b], boxes$lower[b])
# Lanes and labels
plotQTL_lanes(numb_plot)
segments(numb_plot$lane_margins[6,], 0, numb_plot$lane_margins[6,], -numb_plot$track_lengths)
text(numb_plot$lane_centres,     3, 2010:2014, col="gray50", cex=0.7, srt=90)
text(numb_plot$lane_centres[3,], 8, "Italy",   col='red3',   cex=0.8, adj = c(0.5,0))
text(numb_plot$lane_centres[8,], 8, "Sweden",  col='blue4',  cex=0.8, adj = c(0.5,0))

# plot QTL for each site-year combination
for(y in 1:5){
             plotQTL_qtl(numb_plot, numb_stepwise[[y  ]], numb_fit[[y  ]], lane = y,   col = 'red3',  fill_cutoff = 15.2, lwd=2)
  if(y != 4) plotQTL_qtl(numb_plot, numb_stepwise[[y+5]], numb_fit[[y+5]], lane = y+5, col = 'blue4', fill_cutoff = 15.2, lwd=2)
}

# dev.off()
```

Figure S2: QTL for seed number per fruit in each year in Italy (red) and Sweden (blue). Triangle show the maximum-likelihood position of the QTL and bars indicate the extent of 95% Bayesian credible intervals. The direction of the arrow indicates the direction of effect of the Swedish allele on the phenotype. Open triangles are 'poorly defined' QTL with credible intervals greater than 15.2cM.

```{r fig:ind-years-fecundity, fig.width=21/2.56}
mloc <- read.table('../data_files/genetic_map_agren_etal_2013.csv', sep = '\t', header=T)
# pdf('../figures/QTL_positions_tofu.pdf', width=23/2.56, height = 18/2.56, family='Times')
par(mfrow=c(1,1), mar=c(0,3,0,0)+0.1, oma=c(1,0,0,0), family='Times')
# set up plot
tofu_plot <- plotQTL(1:5, mloc, nlanes = 10, left_gap = 4, right_gap = 0)
plotQTL_base(tofu_plot, by=20)
text(33, -92, "Chromosome")
# import fitness boxes and plot.
boxes <- read.csv("../data_files/fitness_boxes.csv")
boxes$track <- match(boxes$chr, colnames(tofu_plot$lane_margins)) # vector denoting which track to plot to.
for(b in 1:nrow(boxes)) plotQTL_box(tofu_plot, boxes$track[b], boxes$upper[b], boxes$lower[b])
# Lanes and labels
plotQTL_lanes(tofu_plot)
segments(tofu_plot$lane_margins[6,], 0, tofu_plot$lane_margins[6,], -tofu_plot$track_lengths)
text(tofu_plot$lane_centres,     3, 2010:2014, col="gray50", cex=0.7, srt=90)
text(tofu_plot$lane_centres[3,], 8, "Italy",   col='red3',   cex=0.8, adj = c(0.5,0))
text(tofu_plot$lane_centres[8,], 8, "Sweden",  col='blue4',  cex=0.8, adj = c(0.5,0))

# plot QTL for each site-year combination
for(y in 1:5){
             plotQTL_qtl(tofu_plot, tofu_stepwise[[y  ]], tofu_fit[[y  ]], lane = y,   col = 'red3',  fill_cutoff = 15.2, lwd=2)
  if(y != 4) plotQTL_qtl(tofu_plot, tofu_stepwise[[y+5]], tofu_fit[[y+5]], lane = y+5, col = 'blue4', fill_cutoff = 15.2, lwd=2)
}
# dev.off()
```

Figure S3: QTL for seed number per plant in each year in Italy (red) and Sweden (blue). Triangle show the maximum-likelihood position of the QTL and bars indicate the extent of 95% Bayesian credible intervals. The direction of the arrow indicates the direction of effect of the Swedish allele on the phenotype. Open triangles are 'poorly defined' QTL with credible intervals greater than 15.2cM.

```{r GLM for effects by year}
load('data_files/qtld.Rdata')
locus_names <- paste("T", sprintf("%02d", 1:19), sep="")

# construct formula for GLMs
form_list <- c("siteyear",
               locus_names,
               paste("siteyear:", locus_names, sep="")
               )
glm_formula <- as.formula(paste("value ~ ", paste(form_list, collapse = " + "), sep=""))

# fit GLMs
reg_byyear <- list(numb = glm(glm_formula, data=qtld$numb),
                   frut = glm(glm_formula, data=qtld$frut, family = 'poisson'),
                   tofu = glm(glm_formula, data=qtld$tofu, family = 'poisson'))

df <- data.frame(sort(unique(qtld$numb$siteyear)),
           matrix(0, nrow=10, ncol=19))
colnames(df) <- c('siteyear', locus_names)

effects_byyear <- list(numb= df, frut = df, tofu = df)

for(t in 1:3){
  ref <- predict(reg_byyear[[t]], df, type = 'response')
  for (i in 1:length(locus_names)){
    this_locus <- df
    this_locus[,locus_names[i]] <- this_locus[,locus_names[i]] + 1
    obs <- predict(reg_byyear[[t]], this_locus, type='response')
    effects_byyear[[t]][,locus_names[i]] <- obs-ref
  }
}
```

```{r fig:effects-by-year, fig.height=17/2.54, fig.width=16.9/2.54}
# setEPS()
# postscript('./figures/effects_by_year.eps', height = 20/2.54, width = 16.9/2.54, family="Times")

par(mfrow=c(3,1), mar=c(3,3,1,1), family="Times")

os <- 0.1

plot(c(1,19), c(-15, 10), type='n', axes = F, xlab="QTL", ylab="")
text(1,10, "A", adj=c(0.5,1))
box(); axis(2);
axis(1, 1:19, paste("T", rownames(clusters$tofu$summary), sep=""), las=2)
title(ylab="Fruit per plant", line = 2)
abline(0,0, lty='dashed')
for(r in 1:5)  points(1:19-os, effects_byyear$frut[r,-1], col='red3', pch=3)
for(r in 6:10) points(1:19+os, effects_byyear$frut[r,-1], col='blue4', pch=3)

plot(c(1,19), c(-5, 5), type='n', axes = F, xlab="QTL", ylab="")
text(1,5, "B", adj=c(0.5,1))
box(); axis(2);
axis(1, 1:19, paste("T", rownames(clusters$tofu$summary), sep=""), las=2)
title(ylab="Seeds per fruit", line = 2)
abline(0,0, lty='dashed')
for(r in 1:5)  points(1:19-os, effects_byyear$numb[r,-1], col='red3', pch=3)
for(r in 6:10) points(1:19+os, effects_byyear$numb[r,-1], col='blue4', pch=3)

plot(c(1,19), c(-600, 400), type='n', axes = F, xlab="QTL", ylab="")
text(1,400, "C", adj=c(0.5,1))
box(); axis(2);
axis(1, 1:19, paste("T", rownames(clusters$tofu$summary), sep=""), las=2)
title(ylab="Seeds per plant", line = 2)
abline(0,0, lty='dashed')
for(r in 1:5)  points(1:19-os, effects_byyear$tofu[r,-1], col='red3', pch=3)
for(r in 6:10) points(1:19+os, effects_byyear$tofu[r,-1], col='blue4', pch=3)
# dev.off()
```

Figure S4: Effect sizes of 19 QTL positions on fruit number per plant, seed number, per fruit and seed number per plant in in each year. Point show effects in Italy (red) and Sweden (blue).