---
title: "Supporting information"
author: "Tom Ellis"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
  bookdown::pdf_book:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, fig.pos = 'H', eval=T)
library("qtl", quietly = T)
library(qtltools)
library(knitr)
```

```{r import-data, eval=T, include=FALSE}
source("R/parents.R")
# get the differences between parental means in each expeirment.
parent_diffs <- data.frame(
  it2010 = parents$delta[parents$site == "italy"  & parents$year == "2010"],
  it2011 = parents$delta[parents$site == "italy"  & parents$year == "2011"],
  sw2010 = parents$delta[parents$site == "sweden" & parents$year == "2010"],
  sw2011 = parents$delta[parents$site == "sweden" & parents$year == "2011"])
parent_diffs <- parent_diffs[c(1,3,2,6,4,7,5),]
parent_diffs$trait <- c("mass", "seed", "frut", "ffit", "surv", "tofu", "tfit")[c(1,3,2,6,4,7,5)]

# import QTL data
source("R/format_QTL_models.R")

```

This Rmarkdown file contains the files necessary to recreate the results presented in the manuscript "Positive correlations among fecundity components boost estimates of local adaptation in *Arabidopsis thaliana*" by Tom Ellis and Jon Ågren. If you are reading a PDF, HTML or MS Word version of this document, see the original `.Rmd` file for the full R source code. See also the file `main_results.Rmd` for code to generate the data in the tables.

```{r cluster-for-table, eval=F}
tabclusters  <- lapply(traits, function(x) return(NULL))
names(tabclusters) <- traits

# perform QTL clustering for each trait
library(qtltools)
for(i in traits){
  if(i == "surv"){
    dat <- cluster_qtl(1:5, qtlmodels[[i]][c(1,2,4)], qtlfits[[i]][c(1,2,4)], 15.2)$full.list
  } else {
    dat <- cluster_qtl(1:5, qtlmodels[[i]], qtlfits[[i]], 15.2)
  }
  tabclusters[[i]] <- dat
  
}

tabclusters$surv$experiment[tabclusters$surv$experiment == 3] <- 4
```

```{r table-of-clusters, eval=F}
# vector of columns to summarise
cmns <- c("ML_bayesint", "effect_sizes", "PVE","LOD")

dat <- tabclusters$mass$full.list

qtl <- unique(dat$QTL_id)

tab <- lapply(1:4, function(x) NULL)
for(e in 1:4){
  thisqtl <- as.data.frame(matrix(NA, nrow=length(qtl), ncol=length(cmns)))
  colnames(thisqtl) <- cmns
  
  for(i in 1:length(qtl)){
    q <- qtl[i]
    qx <- dat$experiment == e & dat$QTL_id ==q
    out <- round(dat[qx,cmns],2)
    if(nrow(out) == 0) out <- rep(NA, 4)
    cat(dim(out), e, q, "\n")
    thisqtl[i, ] <- out
  }
  tab[[e]] <- thisqtl
}
```

\newpage

```{r tab:individual-mass-qtl, eval=T}
site_names  <- c("Italy", "Italy", "Sweden", "Sweden")
year_names  <- c(2010, 2011, 2010, 2011)

# summary tables for each experiment
ix <- "mass"


tab <- list(bayesint_table(qtlmodels[[ix]]$it2010, qtlfits[[ix]]$it2010),
            bayesint_table(qtlmodels[[ix]]$it2011, qtlfits[[ix]]$it2011),
            bayesint_table(qtlmodels[[ix]]$sw2010, qtlfits[[ix]]$sw2010),
            bayesint_table(qtlmodels[[ix]]$sw2011, qtlfits[[ix]]$sw2011))

for(i in 1:4){
  tab[[i]]$site <- rep(site_names[i], nrow(tab[[i]])) # add experiment names
  tab[[i]]$year <- rep(year_names[i], nrow(tab[[i]])) # add experiment names
  tab[[i]]$PDE <- 100*abs(tab[[i]]$effect_sizes) / parent_diffs[parent_diffs$trait == ix, i]
}
tab <- do.call('rbind', tab)
# mash together positions with their CIs
tab$position <- paste(round(tab$ML_bayesint,1),
                      " (", round(tab$min_bayesint, 1), "-", round(tab$max_bayesint, 1), ")", sep="")
# print the table
kable(tab[,c("name", "site", "year","chr","position","LOD", "effect_sizes","PVE", "PDE")], digits = c(0,0,0,0,0,1,3,1,1),
      col.names = c("Name", "Site", "Year","Chr.","Position (cM)","LOD", "Effect size","% var. explained", "% diff. explained"),
      caption="QTL for seed mass identifed in individual experiments. QTL positions are shown with their 95% Bayesian credible intervals. Effect sizes are given as absolute values, as a percentage of variance among RIL means, and as a percentage of the difference between parental means.")
```

\newpage

```{r tab:individual-frut-qtl, eval=T}
# summary tables for fruits/plant
ix <- "frut"
tab <- list(bayesint_table(qtlmodels[[ix]]$it2010, qtlfits[[ix]]$it2010),
            bayesint_table(qtlmodels[[ix]]$it2011, qtlfits[[ix]]$it2011),
            bayesint_table(qtlmodels[[ix]]$sw2010, qtlfits[[ix]]$sw2010),
            bayesint_table(qtlmodels[[ix]]$sw2011, qtlfits[[ix]]$sw2011))

for(i in 1:4){
  tab[[i]]$site <- rep(site_names[i], nrow(tab[[i]])) # add experiment names
  tab[[i]]$year <- rep(year_names[i], nrow(tab[[i]])) # add experiment names
  tab[[i]]$PDE <- 100* abs(tab[[i]]$effect_sizes) / parent_diffs[parent_diffs$trait == ix, i]
}
tab <- do.call('rbind', tab)
# mash together positions with their CIs
tab$position <- paste(round(tab$ML_bayesint,1),
                      " (", round(tab$min_bayesint, 1), "-", round(tab$max_bayesint, 1), ")", sep="")
# print the table
kable(tab[,c("name", "site", "year","chr","position","LOD", "effect_sizes","PVE", "PDE")], digits = c(0,0,0,0,0,1,3,1,1),
      col.names = c("Name", "Site", "Year","Chr.","Position (cM)","LOD", "Effect size","% var. explained", "% diff. explained"),
      caption="QTL for fruit number per reproductive plant identifed in individual experiments. QTL positions are shown with their 95% Bayesian credible intervals. Effect sizes are given as absolute values, as a percentage of variance among RIL means, and as a percentage of the difference between parental means.")
```

\newpage

```{r tab:individual-numb-qtl, eval=T}
# seed number per fruit
ix <- "seed"
tab <- list(bayesint_table(qtlmodels[[ix]]$it2010, qtlfits[[ix]]$it2010),
            bayesint_table(qtlmodels[[ix]]$it2011, qtlfits[[ix]]$it2011),
            bayesint_table(qtlmodels[[ix]]$sw2010, qtlfits[[ix]]$sw2010),
            bayesint_table(qtlmodels[[ix]]$sw2011, qtlfits[[ix]]$sw2011))

for(i in 1:4){
  tab[[i]]$site <- rep(site_names[i], nrow(tab[[i]])) # add experiment names
  tab[[i]]$year <- rep(year_names[i], nrow(tab[[i]])) # add experiment names
  tab[[i]]$PDE <- 100 * abs(tab[[i]]$effect_sizes) / parent_diffs[parent_diffs$trait == ix, i]
}
tab <- do.call('rbind', tab)
# mash together positions with their CIs
tab$position <- paste(round(tab$ML_bayesint,1),
                      " (", round(tab$min_bayesint, 1), "-", round(tab$max_bayesint, 1), ")", sep="")
# print the table
kable(tab[,c("name", "site", "year","chr","position","LOD", "effect_sizes","PVE", "PDE")], digits = c(0,0,0,0,0,1,3,1,1,1),
      col.names = c("Name", "Site", "Year","Chr.","Position (cM)","LOD", "Effect size","% var. explained", "% diff. explained"),
      caption="QTL for seed number per fruit identifed in individual experiments. QTL positions are shown with their 95% Bayesian credible intervals. Effect sizes are given as absolute values, as a percentage of variance among RIL means, and as a percentage of the difference between parental means.")
```

\newpage

```{r tab:individual-tofu-qtl, eval=T}
# total fecundity
ix <- "tofu"
tab <- list(bayesint_table(qtlmodels[[ix]]$it2010, qtlfits[[ix]]$it2010),
            bayesint_table(qtlmodels[[ix]]$it2011, qtlfits[[ix]]$it2011),
            bayesint_table(qtlmodels[[ix]]$sw2010, qtlfits[[ix]]$sw2010),
            bayesint_table(qtlmodels[[ix]]$sw2011, qtlfits[[ix]]$sw2011))

for(i in 1:4){
  tab[[i]]$site <- rep(site_names[i], nrow(tab[[i]])) # add experiment names
  tab[[i]]$year <- rep(year_names[i], nrow(tab[[i]])) # add experiment names
  tab[[i]]$PDE <-100* abs(tab[[i]]$effect_sizes) / parent_diffs[parent_diffs$trait == ix, i]
}
tab <- do.call('rbind', tab)
# mash together positions with their CIs
tab$position <- paste(round(tab$ML_bayesint,1),
                      " (", round(tab$min_bayesint, 1), "-", round(tab$max_bayesint, 1), ")", sep="")
# print the table
kable(tab[,c("name", "site", "year","chr","position","LOD", "effect_sizes","PVE", "PDE")], digits = c(0,0,0,0,0,1,3,1,1),
      col.names = c("Name", "Site", "Year","Chr.","Position (cM)","LOD", "Effect size","% var. explained", "% diff. explained"),         caption = "QTL for seed number per reproductive plant identifed in individual experiments. QTL positions are shown with their 95% Bayesian credible intervals. Effect sizes are given as absolute values, as a percentage of variance among RIL means, and as a percentage of the difference between parental means.")

```

\newpage

```{r tab:individual-tfit-qtl, eval=T}
# seeds/seedling
ix <- "tfit"
tab <- list(bayesint_table(qtlmodels[[ix]]$it2010, qtlfits[[ix]]$it2010),
            bayesint_table(qtlmodels[[ix]]$it2011, qtlfits[[ix]]$it2011),
            bayesint_table(qtlmodels[[ix]]$sw2010, qtlfits[[ix]]$sw2010),
            bayesint_table(qtlmodels[[ix]]$sw2011, qtlfits[[ix]]$sw2011))

for(i in 1:4){
  tab[[i]]$site <- rep(site_names[i], nrow(tab[[i]])) # add experiment names
  tab[[i]]$year <- rep(year_names[i], nrow(tab[[i]])) # add experiment names
  tab[[i]]$PDE <- 100*abs(tab[[i]]$effect_sizes) / parent_diffs[parent_diffs$trait == ix, i]
}
tab <- do.call('rbind', tab)
# mash together positions with their CIs
tab$position <- paste(round(tab$ML_bayesint,1),
                      " (", round(tab$min_bayesint, 1), "-", round(tab$max_bayesint, 1), ")", sep="")
# print the table
kable(tab[,c("name", "site", "year","chr","position","LOD", "effect_sizes","PVE", "PDE")], digits = c(0,0,0,0,0,1,3,1,1),
      col.names = c("Name", "Site", "Year","Chr.","Position (cM)","LOD", "Effect size","% var. explained", "% diff. explained"),
      caption = "QTL for seed number per planted seedling identifed in individual experiments. QTL positions are shown with their 95% Bayesian credible intervals. Effect sizes are given as absolute values, as a percentage of variance among RIL means, and as a percentage of the difference between parental means.")

```

\newpage

```{r tab:individual-ffit-qtl, eval=T}
# Fruit number per seedling
ix <- "ffit"
tab <- list(bayesint_table(qtlmodels[[ix]]$it2010, qtlfits[[ix]]$it2010),
            bayesint_table(qtlmodels[[ix]]$it2011, qtlfits[[ix]]$it2011),
            bayesint_table(qtlmodels[[ix]]$sw2010, qtlfits[[ix]]$sw2010),
            bayesint_table(qtlmodels[[ix]]$sw2011, qtlfits[[ix]]$sw2011))

for(i in 1:4){
  tab[[i]]$site <- rep(site_names[i], nrow(tab[[i]])) # add experiment names
  tab[[i]]$year <- rep(year_names[i], nrow(tab[[i]])) # add experiment names
  tab[[i]]$PDE <- 100*abs(tab[[i]]$effect_sizes) / parent_diffs[parent_diffs$trait == ix, i]
}
tab <- do.call('rbind', tab)
# mash together positions with their CIs
tab$position <- paste(round(tab$ML_bayesint,1),
                      " (", round(tab$min_bayesint, 1), "-", round(tab$max_bayesint, 1), ")", sep="")
# print the table
kable(tab[,c("name", "site", "year","chr","position","LOD", "effect_sizes","PVE", "PDE")], digits = c(0,0,0,0,0,1,3,1,1),
      col.names = c("Name", "Site", "Year","Chr.","Position (cM)","LOD", "Effect size","% var. explained", "% diff. explained"),
      caption = "QTL for fruit number per planted seedling identifed in individual experiments. QTL positions are shown with their 95% Bayesian credible intervals. Effect sizes are given as absolute values, as a percentage of variance among RIL means, and as a percentage of the difference between parental means.")

```

\newpage

```{r tab:individual-surv-qtl, eval=T}
# survival to reproduction
site_names  <- c("Italy", "Italy", "Sweden")
year_names  <- c(2010, 2011, 2011)


ix <- "surv"
tab <- list(bayesint_table(qtlmodels[[ix]]$it2010, qtlfits[[ix]]$it2010),
            bayesint_table(qtlmodels[[ix]]$it2011, qtlfits[[ix]]$it2011),
            #bayesint_table(qtlmodels[[ix]]$sw2010, qtlfits[[ix]]$sw2010),
            bayesint_table(qtlmodels[[ix]]$sw2011, qtlfits[[ix]]$sw2011))

pd <- parent_diffs[parent_diffs$trait == ix,c(1,2,4)]

for(i in 1:3){
  tab[[i]]$site <- rep(site_names[i], nrow(tab[[i]])) # add experiment names
  tab[[i]]$year <- rep(year_names[i], nrow(tab[[i]])) # add experiment names
  tab[[i]]$PDE <- (100* abs(tab[[i]]$effect_sizes)) / as.numeric(pd[i])
}
tab <- do.call('rbind', tab)
# mash together positions with their CIs
tab$position <- paste(round(tab$ML_bayesint,1),
                      " (", round(tab$min_bayesint, 1), "-", round(tab$max_bayesint, 1), ")", sep="")
# print the table
kable(tab[,c("name", "site", "year","chr","position","LOD", "effect_sizes","PVE", "PDE")], digits = c(0,0,0,0,0,1,3,1),
      col.names = c("Name", "Site", "Year","Chr.","Position (cM)","LOD", "Effect size","% var. explained", "% diff. explained"),
      caption = "QTL for survival to reproduction identifed in individual experiments. QTL positions are shown with their 95% Bayesian credible intervals. Effect sizes are given as absolute values, as a percentage of variance among RIL means, and as a percentage of the difference between parental means.")

```

\newpage

```{r tab: epistatic-qtl, eval=T}

# function to extract information about epistatic interactions.
epi_fun <- function(x) {
  d <- summary(x)
  i <- nrow(d$result.drop)
  c(row.names(d$ests)[i+1], d$ests[i, 1], d$result.drop[i, c(3,4)])
}

trait_names <- c("Seed mass", "Fruits/RP", "Seeds/fr", "Seeds/RP", "Fruits/sdl", "Seeds/sdl", "Survival")
site_names  <- c("Italy", "Italy", "Sweden", "Sweden")
year_names  <- c(2010, 2011, 2010, 2011)

epi_tab <- matrix(NA, ncol=8, nrow=0)

for(tr in 1:7){
  for(sy in 1:4){
    if(length(qtlmodels[[tr]][[sy]]) > 0){
      if(length(strsplit(formula(qtlmodels[[tr]][[sy]]), ":")[[1]]) > 1){
        tab <- epi_fun(qtlfits[[tr]][[sy]])
        
        tab <- c(tab, abs(round(100*as.numeric(tab[2]) / parent_diffs[parent_diffs$trait== names(qtlfits)[tr], sy], 1)))
        # print(as.numeric(tab[2]) / parent_diffs[parent_diffs$trait== names(qtlfits)[tr], sy])
        epi_tab <- rbind(epi_tab, c(trait_names[tr], site_names[sy], year_names[sy], tab))
      }
    }
  }
}

colnames(epi_tab) <- c("Trait",   "Site", "Year", "QTL", "Effect size", "LOD", "% var. explained", "% diff. explained")
for(c in 5:7) epi_tab[,c] <- round(as.numeric(as.vector(epi_tab[,c])), 2)

kable(epi_tab, digits = c(0,0,0,0,1,1,1, 1),
      caption = "Pairs of significant epistatic QTL  for fruit number per reproductive plant (Fruits/RP), seed number per fruit (Seeds/fr), seed number per reproductive plant (Seeds/RP), seed mass (Sd mass) and survival to reproduction (Surv). Effect sizes are given as absolute values, as a percentage of variance among RIL means, and as a percentage of the difference between parental means.")
```

\newpage
```{r fitness-qtl, eval=T, fig.cap="QTL for estimates of overall fitness in each site-year combination. Lanes show QTL for fruit number per planted seedling (Fruits/sdl) and seed number per seedling (Seeds/sdl). Arrows indicate most-likely QTL position and the effect of the Swedish genotype (upward: increased phenotype; downward: decreased phenotype) in Italy (red) and Sweden (blue). Vertical bars show the 95% Bayesian credible intervals for QTL position. Open arrows show QTL with credible intervals >15.2cM. Grey boxes indicate the range of colocalising QTL for a single trait across site-year combinations."}
library(qtltools)
mloc <- read.table('data_files/genetic_map_agren_etal_2013.csv', sep = '\t', header=T)

# setEPS()
# postscript('./figures/QTL_survival_fitness.eps', width=25/2.56, height = 16/2.54, family='Times')
# pdf('figures/QTL_fitness.pdf', width=16.9/2.56, height = 12/2.54, family='Times')

par(mfrow=c(1,1), mar=c(1,3,0,0)+0.1, oma=c(0,0,0,0))

pos_plot <- plotQTL(1:5, mloc, nlanes = 8)
plotQTL_base(pos_plot, by=20, maxy = 10)
text(35, -95, "Chromosome")

colvec=c("red3", "red3","blue4","blue4")

pos_counter <- 1
for(t in c("ffit", "tfit")){
  # indicate extent of overlapping QTL with grey boxes.
  boxes <- qtlclusters[[t]]
  boxes$track <- match(boxes$chr, colnames(pos_plot$lane_margins)) # vector denoting which track to plot to.
  boxes <- boxes[boxes$chr %in% 1:5,]
  for(b in 1:nrow(boxes)){
    plotQTL_box(pos_plot, boxes$track[b],
                -boxes$upper[b], -boxes$lower[b],
                left_lane = pos_counter, right_lane = pos_counter+4,
                border = F, col='gray75', lwd=2)
  }
  # plot QTL positions
  for(i in 1:4){
    plotQTL_qtl(pos_plot, qtlmodels[[t]][[i]], qtlfits[[t]][[i]],
                fill_cutoff = 15.2, lane = i+pos_counter-1, col=colvec[i])
  }
  pos_counter <- pos_counter + 4
}

# Lanes dividers
segments(pos_plot$lane_margins[-1,], 0,
         pos_plot$lane_margins[-1,], rep(-pos_plot$track_lengths, each=pos_plot$nlanes),
         lty=c(2,2,2,1), col=c('gray50','gray50','gray50',1))
plotQTL_lanes(pos_plot, lty=0)
# Lane labels
text(pos_plot$lane_margins[c(3,7),], 10, c("Fruits/sdl", "Seeds/sdl"), col=1, cex=0.5, srt=0, adj=c(0.5,0.5))
# text(pos_plot$lane_margins[seq(2,21,2),], 2, c("2010", "2011"), cex=0.7, adj=c(0.5,0.5), srt=90)
text(pos_plot$lane_centres, 2, c("10", "11"), cex=0.5, adj=c(0.5,0.5), srt=0, col='gray30')
text(pos_plot$lane_margins[seq(2,9,2),], 6, c("IT", "SW"), cex=0.6, adj=c(0.5,0.5), srt=0, col=c('red3','blue4'))

colvec=c("red3", "red3","blue4","blue4")

# dev.off()

```

```{r fig: epistatic-mass, fig.width=8/2.54, cache=T, cache.lazy=T, fig.cap="Interaction plots for the epistatic interaction detected for seed mass, showing mean and standard errors for each genotype-by-genotype combination."}
# seed mass
par(mfrow=c(1,1))
effectplot(cross$mass, pheno.col = "sw2010", mname1 = "2@22.0", mname2 = "5@2.6", add.legend = F,
           main = "Sweden 2010", xlab="QTL 5:2.6", ylab="Seed mass (mg)")
legend('topleft', legend = c("QTL 2:22.0","ItIt", "SwSw"), col=c(0,2,4), pch=1, bty="n")
```

```{r fig: epistatic-seed, fig.width=16.9/2.54, fig.height=10/2.54, cache=T, cache.lazy=T, fig.cap="Interaction plots for epistatic interactions detected for seed number per fruit, showing mean and standard errors for each genotype-by-genotype combination."}
# seed number per fruit
par(mfrow=c(1,3))
effectplot(cross$seed, pheno.col = "it2011", mname1 = "1@5.3", mname2 = "5@15.0", add.legend = F,
           main = "Italy 2011", xlab="QTL 5:15.0", ylab="Seed number per fruit")
legend('topright', c("QTL 1:5.3","ItIt", "SwSw"), col=c(0,2,4), pch=1, bty="n")

effectplot(cross$seed, pheno.col = "it2011", mname1 = "1@22.7", mname2 = "5@70.1", add.legend = F,
           main = "Sweden 2010", xlab="QTL 5:70.1", ylab="Seed number per fruit")
legend('topright', c("QTL 1:22.7","ItIt", "SwSw"), col=c(0,2,4), pch=1, bty="n")

effectplot(cross$seed, pheno.col = "it2011", mname1 = "2@22.6", mname2 = "5@2.6", add.legend = F,
           main = "Sweden 2011", xlab="QTL 5:2.6", ylab="Seed number per fruit")
legend('topright', c("QTL 2:22.6","ItIt", "SwSw"), col=c(0,2,4), pch=1, bty="n")
```

```{r fig: epistatic-frut, fig.width=8/2.54, fig.height=10/2.54, cache=T, cache.lazy=T, fig.cap="Interaction plots for epistatic interactions detected for fruit number per reproductive plant, showing mean and standard errors for each genotype-by-genotype combination."}
#fruit number per plant
par(mfrow=c(1,1))
# effectplot(cross$frut, pheno.col = "it2010", mname1 = "5@7.7", mname2 = "5@70.1", add.legend = F,
#            main = "Italy 2010", xlab="QTL 5:70.1", ylab="fruit number per reproductive plant")
# legend(1.4,16, c("QTL 5:7.7","ItIt", "SwSw"), col=c(0,2,4), pch=1, bty="n")

effectplot(cross$frut, pheno.col = "sw2011", mname1 = "1@61.8", mname2 = "3@26.7", add.legend = F, ylim=c(16,30),
           main = "Sweden 2011", xlab="QTL 3:26.7", ylab="Fruit number per reproductive plant")
legend('bottomleft', c("QTL 1:61.8","ItIt", "SwSw"), col=c(0,2,4), pch=1, bty="n")
```

```{r fig: epistatic-ffit, fig.width=16.9/2.54, fig.height=10/2.54, cache=T, cache.lazy=T, fig.cap="Interaction plots for epistatic interactions detected for fruit number per planted seedling, showing mean and standard errors for each genotype-by-genotype combination."}
#fruit number per plant
par(mfrow=c(1,2))
effectplot(cross$ffit, pheno.col = "it2010", mname1 = "4@50.0", mname2 = "5@54.8", add.legend = F,
           main = "Italy 2010", xlab="QTL 5:54.8", ylab="fruit number per reproductive plant")
legend('bottomleft', c("QTL 4:50.0", "ItIt", "SwSw"), col=c(0,2,4), pch=1, bty="n")

effectplot(cross$frut, pheno.col = "sw2011", mname1 = "1@61.8", mname2 = "3@26.7", add.legend = F, ylim=c(16,30),
           main = "Sweden 2011", xlab="QTL 3:26.7", ylab="Fruit number per reproductive plant")
legend('bottomleft', c("QTL 1:61.8","ItIt", "SwSw"), col=c(0,2,4), pch=1, bty="n")
```


```{r fig: epistatic-tofu, fig.width=8/2.54, fig.height=10/2.54, cache=T, cache.lazy=T, fig.cap="Interaction plots for epistatic interactions detected for seed number per reproductive plant, showing mean and standard errors for each genotype-by-genotype combination."}
# seed number per plant
par(mfrow=c(1,1))
# effectplot(tofuqtl, pheno.col = "mean_it2010", mname1 = "3@6.3", mname2 = "5@12.0", add.legend = F,
#            main = "Italy 2010", xlab="QTL 5:12.0", ylab="seed number per reproductive plant")
# legend(1.4,450, c("QTL 3:6.3","ItIt", "SwSw"), col=c(0,2,4), pch=1, bty="n")

effectplot(cross$tofu, pheno.col = "sw2011", mname1 = "1@61.1", mname2 = "3@24.4", add.legend = F, ylim=c(500,1100),
           main = "Sweden 2011", xlab="QTL 3:24.4", ylab="Seed number per reproductive plant")
legend('bottomleft', c("QTL 1.61.1","ItIt", "SwSw"), col=c(0,2,4), pch=1, bty="n")

```

```{r fig: epistatic-tfit, fig.width=16.9/2.54, fig.height=10/2.54, cache=T, cache.lazy=T, fig.cap="Interaction plots for epistatic interactions detected for seed number per planted seedling, showing mean and standard errors for each genotype-by-genotype combination."}
# seed number per plant
par(mfrow=c(1,2))
effectplot(cross$tofu, pheno.col = "it2010", mname1 = "5@43.0", mname2 = "5@74.5", add.legend = F,
           main = "Italy 2010", xlab="QTL 5:74.5", ylab="seed number per reproductive plant")
legend('bottomleft', c("QTL 3:6.3","ItIt", "SwSw"), col=c(0,2,4), pch=1, bty="n")

effectplot(cross$tofu, pheno.col = "sw2011", mname1 = "1@61.1", mname2 = "3@24.4", add.legend = F, ylim=c(500,1100),
           main = "Sweden 2011", xlab="QTL 3:24.4", ylab="Seed number per reproductive plant")
legend('bottomleft', c("QTL 1.61.1","ItIt", "SwSw"), col=c(0,2,4), pch=1, bty="n")

```
