---
title: "Supporting information"
author: "Tom Ellis"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  bookdown::pdf_book:
    toc: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, fig.pos = 'H')
library("qtl", quietly = T)
library(qtltools)
```

This Rmarkdown file contains the files necessary to recreate the results presented in the manuscript "Positive correlations among fecundity components boost estimates of local adaptation in *Arabidopsis thaliana*" by Tom Ellis and Jon Ågren. If you are reading a PDF, HTML or MS Word version of this document, see the original `.Rmd` file for the full R source code. See also the file `main_results.Rmd` for code to generate the data in the tables.

Summary of the genotype data in the R/QTL cross files.

```{r import-qtl-data, include=FALSE}
# Import R/qtl data on seed number, mass and total estimated fitness (seed number * total fruit number).
library("qtl", quietly = T)

# There are QTL models for each of four site-year combinations
siteyear <- c("it2010", "it2011", "sw2010", "sw2011")
# There are six phenotypes in each site-year combination:
traits <- c(
  "mass", # 1. seed mass (mass)
  "frut", # 2. fruits/plant ()
  "seed", # 3. seeds/fruit
  "tofu", # 4. Total fecundity (estimated seeds/plant = frut*seed)
  "ffit", # 5. Fruits per planted seedling; fitness measure used by Agren etal 2013
  "tfit", # 6. Seeds per planted seedling; fitness incorporating seed number.
  "surv") # 7. Survival
# parameters to input
ix <- traits[7] # pull out the name of the trait we are mapping here

# The following loop imports cross and qtl objects, then fits ANOVA models and clusters into groups.
# first, set up empty lists to hold this information.
cross <-              qtlmodels  <-       qtlfits  <- lapply(traits, function(x) return(NULL))
names(cross) <- names(qtlmodels) <- names(qtlfits) <- traits

for(i in traits){
  # Import rqtl cross objects.
  thiscross <- read.cross("csv", "", paste("data_files/rqtl_", i,".csv",sep=""),
                          na.strings=NA, genotypes = c("a", "b"), alleles=c("It","Sw"))
  thiscross <- convert2riself(thiscross)
  thiscross <- calc.genoprob(thiscross, step=2, error.prob=.0001, map.function="kosambi",stepwidth="max", off.end=0)
  cross[[i]] <- thiscross
}

qtlmodels   <- readRDS(file='output/qtl_stepwise_models.rds')
qtlclusters <- readRDS(file='output/qtl_clusters.rds')
qtlfits     <- readRDS(file='output/qtl_model_fits.rds')


```

```{r cluster-for-table, eval=F}
tabclusters  <- lapply(traits, function(x) return(NULL))
names(tabclusters) <- traits

# perform QTL clustering for each trait
library(qtltools)
for(i in traits){
  if(i == "surv"){
    dat <- cluster_qtl(1:5, qtlmodels[[i]][c(1,2,4)], qtlfits[[i]][c(1,2,4)], 15.2)$full.list
  } else {
    dat <- cluster_qtl(1:5, qtlmodels[[i]], qtlfits[[i]], 15.2)
  }
  tabclusters[[i]] <- dat
  
}

tabclusters$surv$experiment[tabclusters$surv$experiment == 3] <- 4
```

```{r table-of-clusters, eval=F}
# vector of columns to summarise
cmns <- c("ML_bayesint", "effect_sizes", "PVE","LOD")

dat <- tabclusters$mass$full.list

qtl <- unique(dat$QTL_id)

tab <- lapply(1:4, function(x) NULL)
for(e in 1:4){
  thisqtl <- as.data.frame(matrix(NA, nrow=length(qtl), ncol=length(cmns)))
  colnames(thisqtl) <- cmns
  
  for(i in 1:length(qtl)){
    q <- qtl[i]
    qx <- dat$experiment == e & dat$QTL_id ==q
    out <- round(dat[qx,cmns],2)
    if(nrow(out) == 0) out <- rep(NA, 4)
    cat(dim(out), e, q, "\n")
    thisqtl[i, ] <- out
  }
  tab[[e]] <- thisqtl
}
```

\newpage

```{r tab:individual-mass-qtl}
siteyear <- c("italy 2010", "italy 2011", "sweden 2010", "sweden 2011")

# summary tables for each experiment
mass_tab <- list(bayesint_table(mass_stepwise$it2010, mass_fit$it2010),
                 bayesint_table(mass_stepwise$it2011, mass_fit$it2011),
                 bayesint_table(mass_stepwise$sw2010, mass_fit$sw2010),
                 bayesint_table(mass_stepwise$sw2011, mass_fit$sw2011))
for(i in 1:4) mass_tab[[i]]$experiment <- rep(siteyear[i], nrow(mass_tab[[i]])) # add experiment names
mass_tab <- do.call('rbind', mass_tab)
# mash together positions with their CIs
mass_tab$position <- paste(round(mass_tab$ML_bayesint,1), " (", round(mass_tab$min_bayesint, 1), "-", round(mass_tab$max_bayesint, 1), ")", sep="")
# print the table
knitr::kable(mass_tab[,c("experiment","chr","position","LOD", "effect_sizes","PVE")], digits = c(0,0,0,1,3,1),
             col.names = c("experiment","chromosome","position (cM)","LOD", "effect size","% var. explained"),
             caption="QTL for seed mass identifed in individual experiments. QTL positions are shown with their 95% Bayesian credible intervals.")
```

\newpage

```{r tab:individual-frut-qtl}
# fruit per plant
frut_tab <- list(bayesint_table(frut_stepwise$it2010, frut_fit$it2010),
                 bayesint_table(frut_stepwise$it2011, frut_fit$it2011),
                 bayesint_table(frut_stepwise$sw2010, frut_fit$sw2010),
                 bayesint_table(frut_stepwise$sw2011, frut_fit$sw2011))
for(i in 1:4) frut_tab[[i]]$experiment <- rep(siteyear[i], nrow(frut_tab[[i]])) # add experiment names
frut_tab <- do.call('rbind', frut_tab)
# mash together positions with their CIs
frut_tab$position <- paste(round(frut_tab$ML_bayesint,1), " (", round(frut_tab$min_bayesint, 1), "-", round(frut_tab$max_bayesint, 1), ")", sep="")
# print the table
knitr::kable(frut_tab[,c("experiment","chr","position","LOD", "effect_sizes","PVE")], digits = c(0,0,0,1,3,1),
             col.names = c("experiment","chromosome","position (cM)","LOD", "effect size","% var. explained"),
             caption="QTL for fruit number per reproductive plant identifed in individual experiments. QTL positions are shown with their 95% Bayesian credible intervals.")
```

\newpage

```{r tab:individual-numb-qtl}
# seed number per fruit
seed_tab <- list(bayesint_table(seed_stepwise$it2010, seed_fit$it2010),
                 bayesint_table(seed_stepwise$it2011, seed_fit$it2011),
                 bayesint_table(seed_stepwise$sw2010, seed_fit$sw2010),
                 bayesint_table(seed_stepwise$sw2011, seed_fit$sw2011))
for(i in 1:4) seed_tab[[i]]$experiment <- rep(siteyear[i], nrow(seed_tab[[i]])) # add experiment names
seed_tab <- do.call('rbind', seed_tab)
# mash together positions with their CIs
seed_tab$position <- paste(round(seed_tab$ML_bayesint,1), " (", round(seed_tab$min_bayesint, 1), "-", round(seed_tab$max_bayesint, 1), ")", sep="")
# print the table
knitr::kable(seed_tab[,c("experiment","chr","position","LOD", "effect_sizes","PVE")], digits = c(0,0,0,1,3,1),
             col.names = c("experiment","chromosome","position (cM)","LOD", "effect size","% var. explained"),
             caption="QTL for seed number per fruit identifed in individual experiments. QTL positions are shown with their 95% Bayesian credible intervals.")
```

\newpage

```{r tab:individual-tofu-qtl}
# total fecundity
tofu_tab <- list(bayesint_table(tofu_stepwise$it2010, tofu_fit$it2010),
                 bayesint_table(tofu_stepwise$it2011, tofu_fit$it2011),
                 bayesint_table(tofu_stepwise$sw2010, tofu_fit$sw2010),
                 bayesint_table(tofu_stepwise$sw2011, tofu_fit$sw2011))
for(i in 1:4) tofu_tab[[i]]$experiment <- rep(siteyear[i], nrow(tofu_tab[[i]])) # add experiment names
tofu_tab <- do.call('rbind', tofu_tab)
# mash together positions with their CIs
tofu_tab$position <- paste(round(tofu_tab$ML_bayesint,1), " (", round(tofu_tab$min_bayesint, 1), "-", round(tofu_tab$max_bayesint, 1), ")", sep="")
# print the table
knitr::kable(tofu_tab[,c("experiment","chr","position","LOD", "effect_sizes","PVE")], digits = c(0,0,0,1,1,1),
             col.names = c("experiment","chromosome","position (cM)","LOD", "effect size","% var. explained"),
             caption = "QTL for seed number per reproductive plant identifed in individual experiments. QTL positions are shown with their 95% Bayesian credible intervals.")

```

```{r tab: epistatic-qtl}

lst <- list(mass_fit$sw2011, frut_fit$it2010, frut_fit$sw2011, 
     seed_fit$it2011, seed_fit$sw2010, seed_fit$sw2011, tofu_fit$it2010, tofu_fit$sw2011)

epi_fun <- function(x) {
  d <- summary(x)
  i <- nrow(d$result.drop)
  c(row.names(d$ests)[i+1], d$ests[i, 1], d$result.drop[i, c(3,4)])
}

epi_tab <- cbind(trait = c("seed mass", 
                           "fruits per reproductive adult", "fruits per reproductive adult", "fruits per reproductive adult",
                           "seeds per fruit", "seeds per fruit",
                           "seeds per reproductive adult", "seeds per reproductive adult"),
                 site = c("sweden", "italy", "sweden", "italy", "sweden", "sweden", "italy", "sweden"),
                 year = c(2011, 2010, 2011, 2011, 2010, 2011, 2010, 2011),
                 as.data.frame(t(sapply(lst, epi_fun)))
)
colnames(epi_tab) <- c("trait",   "site", "year", "loci", "effect size", "LOD", "% variance explained")
for(c in 5:7) epi_tab[,c] <- as.numeric(as.vector(epi_tab[,c]))

knitr::kable(epi_tab, digits = 2,
             caption = "Pairs of significant epistatic QTL.")
rm(lst)
```

\newpage
```{r fitness-qtl, eval=T, fig.cap="QTL for estimates of overall fitness in each site-year combination. Lanes show QTL for fruit number per planted seedling (Fruits/sdl) and seed number per seedling (Seeds/sdl). Arrows indicate most-likely QTL position and the effect of the Swedish genotype (upward: increased phenotype; downward: decreased phenotype) in Italy (red) and Sweden (blue). Vertical bars show the 95% Bayesian credible intervals for QTL position. Open arrows show QTL with credible intervals >15.2cM. Grey boxes indicate the range of colocalising QTL for a single trait across site-year combinations."}
library(qtltools)
mloc <- read.table('data_files/genetic_map_agren_etal_2013.csv', sep = '\t', header=T)

# setEPS()
# postscript('./figures/QTL_survival_fitness.eps', width=25/2.56, height = 16/2.54, family='Times')
# pdf('figures/QTL_fitness.pdf', width=16.9/2.56, height = 12/2.54, family='Times')

par(mfrow=c(1,1), mar=c(1,3,0,0)+0.1, oma=c(0,0,0,0), family='Times')

pos_plot <- plotQTL(1:5, mloc, nlanes = 8)
plotQTL_base(pos_plot, by=20, maxy = 10)
text(35, -95, "Chromosome")

pos_counter <- 1
for(t in c("ffit", "tfit")){
  # indicate extent of overlapping QTL with grey boxes.
  boxes <- qtlclusters[[t]]
  boxes$track <- match(boxes$chr, colnames(pos_plot$lane_margins)) # vector denoting which track to plot to.
  boxes <- boxes[boxes$chr %in% 1:5,]
  for(b in 1:nrow(boxes)){
    plotQTL_box(pos_plot, boxes$track[b],
                -boxes$upper[b], -boxes$lower[b],
                left_lane = pos_counter, right_lane = pos_counter+4,
                border = F, col='gray75', lwd=2)
  }
  # plot QTL positions
  for(i in 1:4){
    plotQTL_qtl(pos_plot, qtlmodels[[t]][[i]], qtlfits[[t]][[i]],
                fill_cutoff = 15.2, lane = i+pos_counter-1, col=colvec[i])
  }
  pos_counter <- pos_counter + 4
}

# Lanes dividers
segments(pos_plot$lane_margins[-1,], 0,
         pos_plot$lane_margins[-1,], rep(-pos_plot$track_lengths, each=pos_plot$nlanes),
         lty=c(2,2,2,1), col=c('gray50','gray50','gray50',1))
plotQTL_lanes(pos_plot, lty=0)
# Lane labels
text(pos_plot$lane_margins[c(3,7),], 8, c("Fruits/sdl", "Seeds/sdl"), col=1, cex=0.6, srt=0, adj=c(0.5,0.5))
# text(pos_plot$lane_margins[seq(2,21,2),], 2, c("2010", "2011"), cex=0.7, adj=c(0.5,0.5), srt=90)
text(pos_plot$lane_centres, 2, c("10", "11"), cex=0.6, adj=c(0.5,0.5), srt=0, col='gray30')
text(pos_plot$lane_margins[seq(2,9,2),], 6, c("IT", "SW"), cex=0.6, adj=c(0.5,0.5), srt=0, col=c('red3','blue4'))

colvec=c("red3", "red3","blue4","blue4")

# dev.off()

```



```{r fig: epistatic-mass, fig.width=8/2.54, cache=T, cache.lazy=T, fig.cap="Interaction plots for the epistatic interaction detected for seed mass, showing mean and standard errors for each genotype x genotype combination."}
# seed mass
par(mfrow=c(1,1))
effectplot(massqtl, pheno.col = "mean_sw2010", mname1 = "2@22.6", mname2 = "5@2.6", add.legend = F,
           main = "Sweden 2010", xlab="QTL 5:2.6", ylab="seed mass (mg)")
legend(0.9,26.8, c("QTL 3:65.9","ItIt", "SwSw"), col=c(0,2,4), pch=1, bty="n")
```

```{r fig: epistatic-seed, fig.width=16.9/2.54, fig.height=10/2.54, cache=T, cache.lazy=T, fig.cap="Interaction plots for epistatic interactions detected for seed number per fruit, showing mean and standard errors for each genotype x genotype combination."}
# seed number per fruit
par(mfrow=c(1,3))
effectplot(seedqtl, pheno.col = "mean_it2011", mname1 = "1@5.3", mname2 = "5@15.0", add.legend = F,
           main = "Italy 2011", xlab="QTL 5:15.0", ylab="seed number per fruit")
legend(1.4,30.5, c("QTL 1:5.3","ItIt", "SwSw"), col=c(0,2,4), pch=1, bty="n")

effectplot(seedqtl, pheno.col = "mean_sw2010", mname1 = "1@22.7", mname2 = "5@70.1", add.legend = F,
           main = "Sweden 2010", xlab="QTL 5:70.1", ylab="seed number per fruit")
legend(1.4,33, c("QTL 1:22.7","ItIt", "SwSw"), col=c(0,2,4), pch=1, bty="n")

effectplot(seedqtl, pheno.col = "mean_sw2011", mname1 = "2@22.6", mname2 = "5@2.6", add.legend = F,
           main = "Sweden 2011", xlab="QTL 5:2.6", ylab="seed number per fruit")
legend(1.4,29.7, c("QTL 2:22.6","ItIt", "SwSw"), col=c(0,2,4), pch=1, bty="n")
```

```{r fig: epistatic-frut, fig.width=16.9/2.54, fig.height=10/2.54, cache=T, cache.lazy=T, fig.cap="Interaction plots for epistatic interactions detected for fruit number per reproductive plant, showing mean and standard errors for each genotype x genotype combination."}
#fruit number per plant
par(mfrow=c(1,2))
effectplot(frutqtl, pheno.col = "mean_it2010", mname1 = "5@7.7", mname2 = "5@70.1", add.legend = F,
           main = "Italy 2010", xlab="QTL 5:70.1", ylab="fruit number per reproductive plant")
legend(1.4,16, c("QTL 5:7.7","ItIt", "SwSw"), col=c(0,2,4), pch=1, bty="n")
effectplot(frutqtl, pheno.col = "mean_sw2011", mname1 = "1@61.8", mname2 = "3@26.7", add.legend = F, ylim=c(16,30),
           main = "Sweden 2011", xlab="QTL 3:26.7", ylab="fruit number per reproductive plant")
legend(0.8,21, c("QTL 1:61.8","ItIt", "SwSw"), col=c(0,2,4), pch=1, bty="n")
```

```{r fig: epistatic-tofu, fig.width=16.9/2.54, fig.height=10/2.54, cache=T, cache.lazy=T, fig.cap="Interaction plots for epistatic interactions detected for seed number per reproductive plant, showing mean and standard errors for each genotype x genotype combination."}
# seed number per plant
par(mfrow=c(1,2))
effectplot(tofuqtl, pheno.col = "mean_it2010", mname1 = "3@6.3", mname2 = "5@12.0", add.legend = F,
           main = "Italy 2010", xlab="QTL 5:12.0", ylab="seed number per reproductive plant")
legend(1.4,450, c("QTL 3:6.3","ItIt", "SwSw"), col=c(0,2,4), pch=1, bty="n")

effectplot(tofuqtl, pheno.col = "mean_sw2011", mname1 = "1@61.1", mname2 = "3@24.4", add.legend = F, ylim=c(500,1100),
           main = "Sweden 2011", xlab="QTL 3:24.4", ylab="seed number per reproductive plant")
legend(0.8,700, c("QTL 1.61.1","ItIt", "SwSw"), col=c(0,2,4), pch=1, bty="n")

```
